---
title: "Covid91 vaccine study Final2020"
author: "Kristina Larson"
date: "November 18, 2020"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode aids to say covid and then remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine.

# guys
```{r}
guysData <- filter(FinalData,sex=="guy")
```
```{r}
barchartGC(~infected + treatment,data=guysData)
barchartGC(~infected + treatment,data=guysData, type="percent")

```
  
Per the graph, the drug is not effective for the guys. It appears to not be helping keep the guys from contracting Covid. More people who took the drug got sick than those who took the placebo.  

### Numerical Summaries for guys 

```{r}
table1 <- xtabs(~infected + treatment, data=guysData)
rowPerc(table1)
colPerc(table1)
```
  
It appears that 58% of the people that got infected took the drug, and 42% had taken the placebo. 

  
## Inferential for guys

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
The odds of a guy getting Covid if he takes the drug are 1.5:1 greater than if he takes the placebo. 

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# gals
```{r}
galsData <- filter(FinalData,sex=="gal")
```
```{r}
barchartGC(~infected + treatment,data=galsData)
barchartGC(~infected + treatment,data=galsData, type="percent")

```
  
It appears that the drug seems t be somewhat effective for the gals. Fewer people got sick when taking the drug than when taking the placebo. Approximately 64% of the people took the placebo got sick, whereas approximately 36% of the people took the drug got sick. 

### Numerical Summaries for gals 

```{r}
table1 <- xtabs(~infected + treatment, data=galsData)
rowPerc(table1)
colPerc(table1)
```
  
The percentage of gals that got infected on the drug was 33% and on the placebo was 67%. 
  
## Inferential for gals

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The odds of getting covid if you are a gal are about .5:1. 


<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

# drug users
```{r}
DrugData <- filter(FinalData,DrugUser=="yes")
```

```{r}
barchartGC(~infected + treatment,data=DrugData)
barchartGC(~infected + treatment,data=DrugData, type="percent")
```

The drug is working about the same for the drug users as it did for the gals. Approximately 64% of those who took placebo got sick, and approximately 36% of those who took the drug got sick. 
  
### Numerical Summaries for Drug Users 

```{r}
table1 <- xtabs(~infected + treatment, data=DrugData)
rowPerc(table1)
colPerc(table1)
```
  
The percentage of drug users that got infected while taking the drug was 33% and while taking the placebo was 67%. 
  
## Inferential for Drug User

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
  
The odds of getting Covid if you are a drug user is about .55:1. 

# lgbtq
```{r}
lgbtqData <- filter(FinalData,LGBTQ=="gay")
```
```{r}
barchartGC(~infected + treatment,data=lgbtqData)
barchartGC(~infected + treatment,data=lgbtqData, type="percent")

```
  
In this graph, it appears that the drug is not at all effective for the LGBTQ community. More than twice the number of people who took the drug than the placebo got sick with Covid.  

  
### Numerical Summaries for LGBTQ 

```{r}
table1 <- xtabs(~infected + treatment, data=lgbtqData)
rowPerc(table1)
colPerc(table1)
```
  
In the LGBTQ community, the percentage of people who got infected on the drug was nearly 71%, whereas on the placebo was only 29%. 
  
## Inferential for lgbtq

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
  
If you belong to the LGBTQ community, the chances of getting infected if taking the drug are 2.5:1. 

  
# Discussion and Conclusion

  
Based on the information available in the inferential section of this report, I would not recommend the drug for the guys that would be seeking this treatment for Covid 91. 

I would be much more likely to recommend the drug to the gals requesting treatment for Covid 91, based on the efficacy indicated in this study.

For the Drug Users that would be coming in to ask me about using this drug for treatment of Covid 91, I would recommend the use of this drug, since it is indicated as being .55:1 ratio of effectiveness. 

For members of the LGBTQ community I would strongly discourage the use of this drug due to the fact that the efficacy ratio is 2.5:1. That is not very good odds at all of it helping if you get infected with Covid 91. 


<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->


